---
title: "Untitled"
author: "Jenny Yijian Huang"
date: "4/1/2022"
output: html_document
---

This script tests out the Markov chain model for Delta using a short simulation study.

# Trying out different hyperparameters for the beta prior.
```{r}
x = seq(0,1,0.05)
plot(x, y = dbeta(x, 1,10))
```


```{r}
test_6 <- run_DAMCMC_complete_h6(
  theta0, cp_Y, N = 20000,
  rho = 0.2, param = "bg", approx = "ldp",
  iota_dist = "exponential", gener = FALSE, b = 1/2,
  thin = 1, plt = 2500,
  par_prior = list(
    a_beta = 1, b_beta = 1,
    a_gamma = 1, b_gamma = 1,
    a_R0 = 2, b_R0 = 2,
    a_lambda = 1, b_lambda = 1,
    a_pi11 = 1, b_pi11 = 10,
    a_pi01 = 1, b_pi01 = 10
  ), 
  # h_hat = c(0.5,0.5,0.5,0.5,0.5), # P(dbeta_i = 1)
  length_delta = 5,
  x_b_ratio = 2,
  gamma=1
  )
```


```{r}
print(analyze_dbeta(test_6$dbeta_lst))

barplot(analyze_dbeta(test_6$dbeta_lst)[[2]], main="Change Point Locations", xlab="Change Point Location",  
 ylab="Proportion of MCMC Samples", names.arg=c("Delta One","Delta Two","Delta Three","Delta Four","Delta Five"), 
 border="blue")
abline(h=1/2, col="blue", lty="dashed", lwd=4)
```

```{r}
# acceptance rate
test_6$theta_rate_accept
```

```{r}
burnin = 1000
library(tidyverse)
sim_df <- create_df_dbeta(test_6$dbeta_lst, test_6$beta_lst, burnin, "X5")
sim_df
```

# Frequencies of the most common configurations
```{r}
sim_df %>% 
  count(delta_beta) %>% 
  mutate(freq = n / sum(n)) %>% 
  arrange(desc(n)) %>% 
  head(10) %>% 
  ggplot(data = ., aes(x = delta_beta, y = freq)) +
  geom_bar(stat = "identity") +
  labs(title = "Frequency of Change Point Configurations", y = "Proportion", x = "Configuration")+ 
  theme(text = element_text(size=10), axis.text.x = element_text(angle = 90))
```
```{r}
sim_t5 <- sim_df %>% 
  count(delta_beta) %>% 
  mutate(freq = n / sum(n)) %>% 
  arrange(desc(n)) %>% 
  head(5)
sim_t5
```

# Compute the mean betas of the top configurations.
```{r}
sim_top_deltabetas <- sim_df %>% 
  filter(delta_beta %in% sim_t5$delta_beta)

sim_mean_betas <- sim_top_deltabetas %>%
  select(delta_beta, X1.y:X6) %>%
  group_by(delta_beta) %>% 
  summarise(across(everything(), mean)) %>% 
  mutate(X7 = X6) # add the stopping week

# flip from cartesian to index. https://tidyr.tidyverse.org/reference/pivot_longer.html
sim_M <- sim_mean_betas %>% 
  pivot_longer(!delta_beta, names_to = "time_unit", values_to = "beta")

#preserve the order of time units.
sim_M$time_unit <- factor(sim_M$time_unit, levels=unique(sim_M$time_unit))
```

```{r}
beta_over_time6 <- sim_top_deltabetas %>% 
  mutate(X7 = X6) %>%  # add the stopping week
  select(ID, X1.y:X7)
```

# Create a spagetti plot
```{r fig.width=2.5, fig.height=2}
beta_over_time6  %>% 
  pivot_longer(cols = -ID,
                 names_to = "time_unit", names_prefix = "V",
                 names_transform = list(Time = as.integer),
                 values_to = "Beta") %>%
    filter(ID %% 100 == 0) %>%
    ggplot(aes(x = time_unit, y = Beta, group = ID)) +
    geom_step(alpha = 0.1, color = "red") +
    labs(title = "Transmission Rate", y = "Beta", x = "Week", color='Change Point Configuration') + 
  scale_x_discrete(labels=seq(0,6,1)) + # add week 6 in.
  expand_limits(y = 0) + geom_hline(yintercept=c(0.0026,.0014,.0036), linetype="dashed", 
                color = "blue")
```

# Plot of posterior mean of the top config.
```{r}
one_config <- sim_M %>% 
  filter(delta_beta == "0_1_0_1_0")
```

```{r}
# Create a line plot of posterior mean for the top configuration (plus true values for beta.)
# Colors
cols <- c("red")

# Basic line plot with points
ggplot(data=one_config, aes(x=time_unit, y=beta, group=delta_beta, color = delta_beta)) +
  geom_step() +
  # geom_line()+
  geom_point() +
  scale_color_manual(values = cols) +
 geom_hline(yintercept=c(0.0026,.0014,.0036), linetype="dashed", 
                color = "blue") +
  labs(title = "Transmission Rate", subtitle="Top Delta", y = "Beta (Mean)", x = "Week", color='Change Point Configuration') + 
  scale_x_discrete(labels=seq(0,6,1)) +
  expand_limits(y = 0)
```



# Traceplots, Posterior densities, ACF plots, and Credible Intervals for each Beta_t.
```{r}
sim_beta_list <- create_mcmc_list(test_6$beta_lst[burnin:20000])
length(sim_beta_list)
```

```{r}
# combine all beta lists into a dataframe.
beta_sim_tidy <- do.call(rbind, Map(data.frame, Iteration = list(c(1:19001)), Beta1 = sim_beta_list[1], Beta2 = sim_beta_list[2], Beta3 = sim_beta_list[3], Beta4 = sim_beta_list[4], Beta5 = sim_beta_list[5], Beta6 = sim_beta_list[6]))
beta_sim_tidy <- beta_sim_tidy %>% 
  remove_burnin(1000)
```

```{r}
# Posterior means
means <- beta_sim_tidy %>%
  dplyr::select(- .data$Iteration) %>%
  {tibble::tibble(
    var  = colnames(.),
    mean = colMeans(.),
  )}

# ESS
ESS = coda::effectiveSize(coda::mcmc(beta_sim_tidy))
ESS # ESS is higher when we correctly specify the model (eg allow for change points.)
```

```{r}
# plot results in a format similar to PDSIR code.
vars <- c("Beta1", "Beta2","Beta3","Beta4","Beta5","Beta6")

for(var in vars)
  print(draw_tp(beta_sim_tidy, var)) # burn in.
for(var in vars)
  print(draw_histogram(beta_sim_tidy, var))
for(var in vars)
  print(draw_acf(beta_sim_tidy, var, 200))
```

beta_1_MLE  beta_2_MLE  beta_3_MLE 
0.002645833 0.001425511 0.003639395

## Complete D-A MCMC with Markov Chain

```{r}
run_DAMCMC_complete_h6 <- function(
  theta, Y, N = 20000,
  rho = 0.5, param = "bg", approx = "ldp",
  iota_dist = "exponential", gener = FALSE, b = 1/2,
  thin = 100, plt = 1000,
  par_prior = list(
    a_beta = 0.01, b_beta = 0.01,
    a_gamma = 1, b_gamma = 1,
    a_R0 = 2, b_R0 = 2,
    a_lambda = 1, b_lambda = 1,
    a_pi11 = 1, b_pi11 = 10,
    a_pi01 = 1, b_pi01 = 10
  ),
  length_delta = 5,
  x_b_ratio = 5,
  gamma=0.1
) {
  
    # Setup
    pi01_save <- pi11_save <- f_save <- b_f_save <- numeric(N / thin)
    accept <- numeric(N)
    b_accept <- numeric(N)
    ss_save <- vector(mode = "list", length = x_b_ratio*N / thin)
    dbeta_save <- beta_save <- vector(mode = "list", length = N / thin)

    
    # init proposal
    # Propose the initial latent path using Y and initial theta.
    SS_current <- list(compatible = FALSE)
    while(! SS_current[["compatible"]]) {
        x_current <- rprop_x(theta, Y, gener, b, iota_dist, approx, NULL, 1) # original D-A-MCMC code.
        SS_current <- suff_stat(x_current, Y, gener, b)
    }
  
    # Plot initial proposal
    x_current_plt <- suff_stat(x_current, Y, gener, b, return_SI = TRUE)
    plot(x = x_current_plt$t, y = x_current_plt$I, main="Initial Latent Epidemic Trajectory",
        xlab="Time", ylab="I(t)")

    # PI: Propose Pi
    pi01_current <- rbeta(1, par_prior['a_pi01'][[1]], par_prior['b_pi01'][[1]])
    pi11_current <- rbeta(1, par_prior['a_pi11'][[1]], par_prior['b_pi11'][[1]])
    
    # 1. Propose dbeta
    dbeta_current <- draw_mm_delta(pi01_current, pi11_current, length_delta)
    dbeta_save[1] <- list(dbeta_current)

    # 2. determine the change-point locations.
    segment_start <- cp_locations(dbeta_current, Y)$segment_start
    segment_end <- cp_locations(dbeta_current, Y)$segment_end
    
    
    # 3. calculate unsegmented SS (Later: function to segment the suffstats.)
    SS_current <- suff_stat(x_current, Y, gener, b, return_SI = FALSE)
    # 3. calculate sufficient statistics for each segment of x_current using the change-point locations calculated.
    segmented_SS_current <- suff_stat2(segment_start, segment_end, x_current, Y, gener, b, return_SI = FALSE)
    

    # 4. Propose (beta | dbeta)
    beta_current <- draw_beta(dbeta_current, segmented_SS_current, iota_dist, gamma, par_prior, param, Y)
    beta_save[1] <- list(beta_current)
    unique_beta_current <- unique(beta_current)
    
    # 5. Calculate current log likelihood
    b_target_current     <- 0
    for(i in 1:length(unique_beta_current)){
      b_target_current <- b_target_current + loglik_infec(unique_beta_current[i], segmented_SS_current[[i]], gener, b, iota_dist = "exponential")
    }
    b_target_current <- b_target_current + loglik_remov_infectious(gamma, SS_current, gener, b, iota_dist = "exponential")


    # MH 
    for(s in 2 : N) {
      
      # Propose (dbeta, beta | Z) ###########
      
      # Pi Gibbs Update:
      pi01_current <- gibbs_pi(par_prior,dbeta_current)[1]
      pi11_current <- gibbs_pi(par_prior,dbeta_current)[2]
    
      # 1. Propose dbeta
      dbeta_new <- draw_mm_delta(pi01_current, pi11_current, length_delta)
      # dbeta_new <- draw_dbeta(dbeta_current, is_prob=TRUE, Pi_new)[[1]]
      # u_new <- draw_dbeta(dbeta_current, is_prob=TRUE, Pi_new)[[2]]

      # 2. determine the change-point locations.
      segment_start <- cp_locations(dbeta_new, Y)$segment_start
      segment_end <- cp_locations(dbeta_new, Y)$segment_end
      
      # 3. calculate sufficient statistics for Z using the start and end points.
      segmented_SS_new <- suff_stat2(segment_start, segment_end, x_current, Y, gener, b, return_SI = FALSE) # segmented_SS_new because of change points.
  
      # 4. Propose (beta | dbeta)
      beta_new <- draw_beta(dbeta_new, segmented_SS_new, iota_dist, gamma, par_prior, param, Y)
      unique_beta_current <- unique(beta_current)
      unique_beta_new <- unique(beta_new)

      # 5. Compute the Ratio.
      
      # Target density
      b_target_current     <- 0
      for(i in 1:length(unique_beta_current)){
        b_target_current <- b_target_current + loglik_infec(unique_beta_current[i], segmented_SS_current[[i]], gener, b, iota_dist = "exponential")
      }
      b_target_current <- b_target_current + loglik_remov_infectious(gamma, SS_current, gener, b, iota_dist = "exponential")
      
      b_target_new     <- 0
      for(i in 1:length(unique_beta_new)){
        b_target_new <- b_target_new + loglik_infec(unique_beta_new[i], segmented_SS_new[[i]], gener, b, iota_dist = "exponential")
      }
      b_target_new <- b_target_new + loglik_remov_infectious(gamma, SS_current, gener, b, iota_dist = "exponential")
      
      # Proposal density 
      b_prop_current <- f_delta_mm(dbeta_current, pi01_current, pi11_current) + f_beta_func(unique_beta_current, par_prior, segmented_SS_current)
      b_prop_new <- f_delta_mm(dbeta_new, pi01_current, pi11_current) + f_beta_func(unique_beta_new, par_prior, segmented_SS_new)
    
      # Priors
      b_prior_current <- prior_beta(beta_current, par_prior)
      b_prior_new <- prior_beta(beta_new, par_prior)
      
      # MH ratio
      b_R_log <- b_prior_new - b_prior_current + b_target_new - b_target_current + b_prop_current - b_prop_new
      b_R         <- min(1, exp(b_R_log))
      b_accept[s] <- stats::runif(1) < b_R
      
      ### Debugging Inf values
      if(is.na(b_accept[s]) ){
        print(list("b_target_new" = b_target_new, "b_prior_new" = b_prior_new, "b_prop_new" = b_prop_new,
                    "f_pi_new" = f_pi_func(Pi_new, dbeta_new, par_prior), "f_delta_new" = f_dbeta_func(dbeta_new, Pi_new, u_new), 
                   "f_beta_new" = f_beta_func(unique_beta_new, par_prior, segmented_SS_new), 
                   "prior_pi_new" = prior_pi(Pi_new, par_prior), "prior_delta_new" = prior_dbeta(Pi_new, dbeta_new),"prior_beta_new" = prior_beta(beta_new, par_prior)))
      }
      ###
      
      if(s %% 1000 == 0){
          print("Acceptance Ratio (Beta given data)")
          print(b_R)
      }
      
      
      # Accept/reject new draws
      if(b_accept[s]) {
          # update to values.
          dbeta_current <- dbeta_new
          beta_current        <- beta_new
          b_target_current <- b_target_new
          segmented_SS_current <- segmented_SS_new 
      }
      ##############
      
      ##############
      # Propose (x | dbeta, beta) ######## 
      
      # 1. propose x_new
      x_new    <- b_rprop_x(beta_current, theta, Y, gener, b, iota_dist, approx, x_current, rho)
      i_update <- x_new[["i_update"]]
      
      # 2. the current change-point locations.
      segment_start <- cp_locations(dbeta_current, Y)$segment_start
      segment_end <- cp_locations(dbeta_current, Y)$segment_end
      
      # 3. calculate unsegmented sufficient statistics for x_new
      SS_new <- suff_stat(x_new, Y, gener, b, return_SI = FALSE)
      # 3. calculate sufficient statistics for x_new using the start and end points.
      segmented_SS_new_x <- suff_stat2(segment_start, segment_end, x_new, Y, gener, b, return_SI = FALSE)

      # Target density
      unique_beta_current <- unique(beta_current)
      
      f_target_current <- b_target_current # (beta_current, x_current)

      
      f_target_new     <- 0 # (beta_current, x_new)
      for(t in 1:length(unique_beta_current)){
        f_target_new <- f_target_new + loglik_infec(unique_beta_current[t], segmented_SS_new_x[[t]], gener, b, iota_dist = "exponential")
      }
      f_target_new <- f_target_new + loglik_remov_infectious(gamma, SS_new, gener, b, iota_dist = "exponential")
      
      
  
      # Proposal density
      f_prop_current <- b_dprop_x(beta_current, theta, Y, x_current, i_update, gener, b, iota_dist, approx) # Infinity
      f_prop_new     <- b_dprop_x(beta_current, theta, Y, x_new    , i_update, gener, b, iota_dist, approx)
  
      # MH ratio
      R_log     <- f_target_new - f_target_current - f_prop_new + f_prop_current
      
      ### Debugging Inf values
      if(is.nan(accept[s])){
        print(list("f_target_new" = f_target_new, "f_target_current" = f_target_current, "f_prop_new" = f_prop_new, "f_prop_current" = f_prop_current))
      }
      ###
      
      R         <- min(1, exp(R_log))

      # print(R) # NAN
      
      accept[s] <- stats::runif(1) < R # how do i keep track of the new indexing for the accept vector?
  
      # Accept/reject new draws
      if(accept[s]) {
        x_current        <- x_new
        SS_current       <- SS_new
        f_target_current <- f_target_new
        segmented_SS_current <- segmented_SS_new_x
        
      } # end-if
          
      
      
      ##############
    
      # Plot x accepted.
      if(s %% plt == 0){ # plot the first proposal x.
        x_current_plt <- suff_stat(x_current, Y, gener, b, return_SI = TRUE)
        plot(x = x_current_plt$t, y = x_current_plt$I, main="Latent Epidemic Trajectory",
            xlab="Time", ylab="I(t)")
        
      }
      
      # Save thinned draws for output.
      if(s %% thin == 0){
        j <- s / thin
        
        pi01_save[[j]] <- pi01_current
        pi11_save[[j]] <- pi11_current
        
        f_save[[j]] <- f_target_current
        b_f_save[[j]] <- b_target_current
        
        dbeta_save[j] <- list(dbeta_current)
        beta_save[j] <- list(beta_current)
        
        ss_save[j] <- list("Iteration"=segmented_SS_current) # already a list of lists
        
      }

    } # end-for

  # Output
  out <- list(
    segmented_SS = ss_save, pi01_lst = pi01_save, pi11_lst = pi11_save, dbeta_lst = dbeta_save, beta_lst = beta_save, beta_loglik = b_f_save, x_loglik = f_save, theta_rate_accept = mean(b_accept), x_rate_accept = mean(accept)
  )
  
  return(out)
  }
```