---
title: "duke COVID cases"
author: "Jenny Yijian Huang"
date: "12/22/2021"
output: html_document
---

A: The stochastic SIR takes into account the s compartment, so when susceptible individuals have not been sufficiently depleted, the model can attribute the drop in I(t) to a change in transmission rate (due to policy mandates etc.).

Hence, a change point detection model is useful because based on visualizations of the trajectory alone, it is incredibly difficult to pin-point when the transmission rate changed. The time-varying SIR model can help us to locate and quantify these change-points.

https://coronavirus.duke.edu/covid-testing/spring-2021-duke-covid-testing-tracker/ Duke COVID-19 Incidence Data Spr. 2021.

Q. How effective was the stay-in-place order Duke implemented March 13th, 2021? Right at the end of week 9 (March 8-14th). By how much did beta decrease?


```{r}
# ts
duke_Y2 <- list("T_k" = c(49, 62, 54, 40, 13, 11, 22, 53, 218, 61, 25, 15, 29, 21, 0), "I0" = 34, "S0" = 25000, "ts" = seq(0, 15, by = 1), 
"t_end" = 15) # remove the first and last points.
```


```{r}
plot(x = duke_Y2$ts, y = c(0, duke_Y2$T_k), main="Duke COVID-19 Spr. 2021", sub="Jan. 11-17th to Apr. 19-25th", xlab="Time (each point represents a week)", ylab="New Infection Counts")
```

```{r}
dk_observations6 <- run_DAMCMC_complete_old(
  theta0, duke_Y2, N = 20000,
  rho = 0.2, param = "bg", approx = "ldp",
  iota_dist = "exponential", gener = FALSE, b = 1/2,
  thin = 1, plt = 5000,
  par_prior = list(
    a_beta = 1, b_beta = 1, # are cp locations sensitive to abeta prior?
    a_gamma = 1, b_gamma = 1,
    a_R0 = 2, b_R0 = 2,
    a_lambda = 1, b_lambda = 1
  ), h_hat = c(0.5,0.5,0.5,0.5,0.5,0.5,0.5,0.5,0.5,0.5,0.5,0.5,0.5,0.5), # P(dbeta_i = 1)
  x_b_ratio = 2,
  gamma=1
) 
```

```{r}
length(c(0.5,0.5,0.5,0.5,0.5,0.5,0.5,0.5,0.5,0.5,0.5,0.5,0.5,0.5))
```



```{r}
analyze_dbeta(dk_observations6$dbeta_lst)

barplot(analyze_dbeta(dk_observations6$dbeta_lst)[[2]], main="Delta Configurations", xlab="Change Point Location",  
 ylab="frequencyy", 
 border="blue")
abline(h=0.5, col="blue", lty="dashed", lwd=4)
```

```{r}
dk_observations6$theta_rate_accept # 49%
```

Plot results
```{r}
plot(x = duke_Y2$ts[-1]-0.5, y = duke_Y2$T_k, main="Duke COVID-19 Spr. 2021", sub="Jan. 11-17th to Apr. 19-25th", xlab="Time (each point represents a week)", ylab="New Infection Counts") # the points are plotted in the middle of week (i.e. week 1 results located at 0.5)
abline(v=c(6,9),lwd=2,col="red")
```

# Create Dataframe from raw dbeta and beta lists for frequency plot of configurations.
```{r}
create_df_dbeta <- function(mcmc_output_dbeta, mcmc_output_beta, burnin, strlastX) {
  # strlastX: string value of the last dbeta (change point) index.
  temp_df <- mcmc_output_dbeta[burnin:length(mcmc_output_dbeta)]
  temp_df <- data.frame(t(sapply(temp_df,c)))
  dbeta_lst_df <- temp_df %>% 
    mutate(ID = burnin:length(mcmc_output_dbeta)) %>% 
    unite("delta_beta", X1:strlastX, remove=FALSE)

  beta_lst_raw <- mcmc_output_beta[burnin:length(mcmc_output_beta)]
  beta_lst_df <- data.frame(t(sapply(beta_lst_raw,c)))
  beta_lst_df <- beta_lst_df %>% 
    mutate(ID = burnin:length(mcmc_output_beta))
  
  complete_tibble <- merge(dbeta_lst_df, beta_lst_df, by="ID")

  return(complete_tibble)
}

covid_df <- create_df_dbeta(dk_observations6$dbeta_lst, dk_observations6$beta_lst, burnin, "X14")
```

# Frequencies of the most common configurations
```{r}
covid_df %>% 
  count(delta_beta) %>% 
  mutate(freq = n / sum(n)) %>% 
  arrange(desc(n)) %>% 
  head(10) %>% 
  ggplot(data = ., aes(x = delta_beta, y = freq)) +
  geom_bar(stat = "identity") +
  labs(title = "Frequency of Change Point Configurations", y = "Frequency", x = "Configuration")+ 
  theme(text = element_text(size=10), axis.text.x = element_text(angle = 90))
```

```{r}
covid_t5 <- covid_df %>% 
  count(delta_beta) %>% 
  mutate(freq = n / sum(n)) %>% 
  arrange(desc(n)) %>% 
  head(5)
covid_t5
```

```{r}
# Colors
cols <- c("red", "orange", "grey", "yellow", "grey") # INDEXING ISSUE.

# Basic line plot with points
ggplot(data=covid_M, aes(x=time_unit, y=beta, group=delta_beta, color = delta_beta)) +
  geom_step() +
  # geom_line()+
  geom_point() +
  scale_color_manual(values = cols) +
  labs(title = "Transmission Rate of COVID-19", subtitle="Top 5 Configurations", y = "transmission rate (mean)", x = "Week", color='Change Point Configuration') + 
  scale_x_discrete(labels=seq(0,15,1))
```

# Compute the mean betas of the top configurations.
```{r}
covid_top_deltabetas <- covid_df %>% 
  filter(delta_beta %in% covid_t5$delta_beta)

covid_mean_betas <- covid_top_deltabetas %>%
  select(delta_beta, X1.y:X15) %>%
  group_by(delta_beta) %>% 
  summarise(across(everything(), mean)) %>% 
  mutate(X16 = X15) # add the stopping week

# flip from cartesian to index. https://tidyr.tidyverse.org/reference/pivot_longer.html
covid_M <- covid_mean_betas %>% 
  pivot_longer(!delta_beta, names_to = "time_unit", values_to = "beta")

#preserve the order of time units.
covid_M$time_unit <- factor(covid_M$time_unit, levels=unique(covid_M$time_unit))
```

```{r}
covidbeta_over_time6 <- covid_top_deltabetas %>% 
  mutate(X16 = X15) %>%  # add the stopping week
  select(ID, X1.y:X16)
```

```{r fig.width=2.5, fig.height=2.5}
covidbeta_over_time6 <- covidbeta_over_time6  %>% 
  pivot_longer(cols = -ID,
                 names_to = "time_unit", names_prefix = "V",
                 # names_transform = list(Time = as.integer),
                 values_to = "Beta") %>%
    filter(ID %% 100 == 0)

#preserve the order of time units.
covidbeta_over_time6$time_unit <- factor(covidbeta_over_time6$time_unit, levels=unique(covidbeta_over_time6$time_unit))

covidbeta_over_time6 %>% 
    ggplot(aes(x = time_unit, y = Beta, group = ID)) +
    geom_step(alpha = 0.1, color = "red") +
    labs(title = "Transmission Rate", y = "Beta", x = "Week", color='Change Point Configuration') + 
  scale_x_discrete(labels=seq(0,16,1)) + # add week 6 in.
  expand_limits(y = 0)
```


```{r}
covidbeta_over_time6 %>% 
    ggplot(aes(x = time_unit, y = Beta*25000, group = ID)) +
    geom_step(alpha = 0.1, color = "red") +
    labs(title = "SARS-COV-2 Transmission Rate Falls Following Duke's Stay-in-place Order", subtitle = "Stay-in-place order implemented at the EOW 9", y = "Rt", x = "Week", color='Change Point Configuration') + 
  scale_x_discrete(labels=seq(0,16,1)) + # add week 6 in.
  expand_limits(y = 0)
```

# Create dataframe
```{r}
dk_beta_list <- create_mcmc_list(dk_observations6$beta_lst[burnin:20000])
length(dk_beta_list)
```

```{r}
# combine all beta lists into a dataframe.
beta_dk_tidy <- do.call(rbind, Map(data.frame, Iteration = list(c(1:19001)), Beta6 = dk_beta_list[6], Beta7 = dk_beta_list[7], Beta8 = dk_beta_list[8], Beta9 = dk_beta_list[9], Beta10 = dk_beta_list[10]))
beta_dk_tidy <- beta_dk_tidy %>% 
  remove_burnin(1000)
beta_dk_tidy
```


```{r}
# Posterior means
means <- beta_dk_tidy %>%
  dplyr::select(- .data$Iteration) %>%
  {tibble::tibble(
    var  = colnames(.),
    mean = colMeans(.),
  )}

ESS = coda::effectiveSize(coda::mcmc(beta_dk_tidy))
ESS
```
ESS: [80, 200]

```{r}
# chains mixing very well for uniform proposal.
vars <- c("Beta6", "Beta7","Beta8","Beta9")

for(var in vars)
  print(draw_tp(beta_dk_tidy, var)) # burn in.
for(var in vars)
  print(draw_histogram(beta_dk_tidy, var))
for(var in vars)
  print(draw_acf(beta_dk_tidy, var, 200))
```


```{r}
upper95 <- beta_dk_tidy %>%
  summarise_at(2:6, ~ quantile(., probs = c(0.975))) %>%
  gather("var", "q97.5")

lower95 <- beta_dk_tidy %>%
  summarise_at(2:6, ~ quantile(., probs = c(0.025))) %>%
  gather("var", "q2.5")

CI <- inner_join(lower95, upper95, by = "var")
posterior_df <- inner_join(means, CI, by = "var")
posterior_df
```

Week 6 to 7 saw a 2 fold increase in Beta;
Week 9 to 10 saw a 4.5 fold decrease in Beta.

r0 = Beta * S0 / gamma = Beta * 25000 / gamma
R0_6 = 1.001597
R0_9 = 2.216309
R0_10 = 0.4972857
