---
title: "Simulation_study pt 1"
author: "Jenny Yijian Huang"
date: "1/23/2022"
output: html_document
---

Simulation Study:

# First, we simulate an epidemic with change points in R0 at weeks 2,4 using Gillespie's algorithm.
```{r}
change_point_data <- sim_changepoint_sem(
  S0 = 1e3, I0 = 1e1, t_end = 9,
  theta = list(R0 = c(2.5, 1.5, 3.5), gamma = 1, lambda = 1, shape = 1),
  change_day = c(2,4),
  iota_dist = "exponential",
  gener = FALSE, b = 1/2,
  E0 = 0, type = "SIR" 
)
cp_data <- change_point_data$x

print(change_point_data$MLE)

plot(x = change_point_data$t[1:2000], y = change_point_data$I[1:2000], main="Epidemic Trajectory",
   xlab="Time", ylab="I(t)")
```

# Below, we visualize the I(t) compartment over time.
```{r}
plot(x = change_point_data$t[1:2000], y = change_point_data$I[1:2000], main="Epidemic Trajectory",
   xlab="Time", ylab="I(t)")
```


beta_i = governs the time unit from [i, i+1].
delta_i = marks a change point located at i.

# We will use only the partially observed (incidence) data as input to our DA-MCMC.
# (LATER) vary observation schedule.
```{r}
cp_Y <- observed_data(change_point_data, K=10)
plot(x = cp_Y$ts, y = c(0, cp_Y$T_k), main="Observed Counts", xlab="Time", ylab="New Infections")
```

# We see that the chain still converges to the same estimate if we initialize beta to start at a poor initial value.
```{r}
theta0 = list(R0 = 1, beta=0.001, lambda = 1, shape = 1, gamma = 1)
```

```{r}
test_6 <- run_DAMCMC_complete_smallh6(
  theta0, cp_Y, N = 19000,
  rho = 0.2, param = "bg", approx = "ldp",
  iota_dist = "exponential", gener = FALSE, b = 1/2,
  thin = 1, plt = 2500,
  par_prior = list(
    a_beta = 0.1, b_beta = 0.1,
    a_gamma = 1, b_gamma = 1,
    a_R0 = 2, b_R0 = 2,
    a_lambda = 1, b_lambda = 1,
    a_pi11 = .5, b_pi11 = .5,
    a_pi01 = .5, b_pi01 = .5
  ), 
  length_delta = 8,
  x_b_ratio = 2,
  gamma=1
  )
```

```{r}
print(analyze_dbeta(test_6$dbeta_lst))

barplot(analyze_dbeta(test_6$dbeta_lst)[[2]], main="Delta Configurations", xlab="Change Point Location",  
 ylab="Count", names.arg=c("Delta One","Delta Two","Delta Three","Delta Four","Delta Five","Delta 6","Delta 7","Delta 8"), 
 border="blue")
abline(h=1/2, col="blue", lty="dashed", lwd=4)
```

```{r}
burnin = 1000
library(tidyverse)
sim_df <- create_df_dbeta(test_6$dbeta_lst, test_6$beta_lst, burnin, "X8")
sim_df
```
# Frequencies of the most common configurations
```{r}
sim_df %>% 
  count(delta_beta) %>% 
  mutate(freq = n / sum(n)) %>% 
  arrange(desc(n)) %>% 
  head(10) %>% 
  ggplot(data = ., aes(x = delta_beta, y = freq)) +
  geom_bar(stat = "identity") +
  labs(title = "Frequency of Change Point Configurations", y = "Frequency", x = "Configuration")+ 
  theme(text = element_text(size=10), axis.text.x = element_text(angle = 90))
```

```{r}
sim_t5 <- sim_df %>% 
  count(delta_beta) %>% 
  mutate(freq = n / sum(n)) %>% 
  arrange(desc(n))
sim_t5
```


# Compute the mean betas of the top configurations.
```{r}
sim_top_deltabetas <- sim_df %>% 
  filter(delta_beta %in% sim_t5$delta_beta)

sim_mean_betas <- sim_top_deltabetas %>%
  select(ID, delta_beta, X1.y:X9) %>%
  group_by(delta_beta) %>% 
  summarise(across(everything(), mean)) %>% 
  mutate(X10 = X9)

# flip from cartesian to index. https://tidyr.tidyverse.org/reference/pivot_longer.html
sim_M <- sim_mean_betas %>% 
  pivot_longer(!delta_beta, names_to = "time_unit", values_to = "beta")

#preserve the order of time units.
sim_M$time_unit <- factor(sim_M$time_unit, levels=unique(sim_M$time_unit))
```

```{r}
# mutate(r0)
sim_M <- sim_M %>% 
  mutate(R0 = beta * 1000)

# spaghetti plot
# spagetti plot of R0 over time
sim_M %>% 
    ggplot(aes(x = time_unit, y = beta, group=ID)) +
    geom_step(alpha = 0.1, color = "red") +
    labs(title = "Transmission Rate Over time", y = "Beta(t)", x = "Time Units", color='Change Point Configuration') + 
  scale_x_discrete(labels=seq(1,10,1)) +
  #expand_limits(y = 0) +
  theme_bw()
```

```{r}
plot(x = cp_Y$ts, y = c(0, cp_Y$T_k), main="Observed Counts", xlab="Week", ylab="New Infections")
abline(v=c(2,4),lwd=2,col="red")
```


# Create dataframe
```{r}
sim_beta_list <- create_mcmc_list(test_6$beta_lst[burnin:20000])
length(sim_beta_list[[1]])
```

```{r}
# combine all beta lists into a dataframe.
beta_sim_tidy <- do.call(rbind, Map(data.frame, Iteration = list(c(1:18001)), Beta1 = sim_beta_list[1], Beta2 = sim_beta_list[2], Beta3 = sim_beta_list[3], Beta4 = sim_beta_list[4], Beta5 = sim_beta_list[5], Beta6 = sim_beta_list[6],Beta7 = sim_beta_list[7],Beta8 = sim_beta_list[8],Beta9 = sim_beta_list[9]))
```

```{r}
# Posterior means
means <- beta_sim_tidy %>%
  dplyr::select(- .data$Iteration) %>%
  {tibble::tibble(
    var  = colnames(.),
    mean = colMeans(.),
  )}

# ESS
ESS = coda::effectiveSize(coda::mcmc(beta_sim_tidy))
ESS # ESS is higher when we correctly specify the model (eg allow for change points.) ()
```

```{r}
upper95 <- beta_sim_tidy %>%
  summarise_at(2:7, ~ quantile(., probs = c(0.975))) %>%
  gather("var", "q97.5")

lower95 <- beta_sim_tidy %>%
  summarise_at(2:7, ~ quantile(., probs = c(0.025))) %>%
  gather("var", "q2.5")

CI <- inner_join(lower95, upper95, by = "var")
posterior_df <- inner_join(means, CI, by = "var")
posterior_df
```

```{r}
# plot results in a format similar to PDSIR code.
vars <- c("Beta2", "Beta4","Beta6","Beta8")

for(var in vars)
  print(draw_tp(beta_sim_tidy, var, thin = 10)) # burn in.
for(var in vars)
  print(draw_histogram(beta_sim_tidy, var))
for(var in vars)
  print(draw_acf(beta_sim_tidy, var, 200))
```

