---
title: 'ebola data'
author: "Jenny Yijian Huang"
date: "2/11/2022"
output: html_document
---

# Ebola incidence data.
Simulating data from the parameter results.

# Reading in data
```{r}
path_data = "Ebola/Guinea_modified.csv"
```

```{r}
# Ebola Data
d <- readr::read_csv(
  path_data,
  col_types = readr::cols(
    .default            = readr::col_double(),
    Location            = readr::col_character(),
    `Ebola data source` = readr::col_character(),
    `Indicator type`    = readr::col_character(),
    `Case definition`   = readr::col_character()
  )
) %>%
  dplyr::select(-(2:4)) %>%
  dplyr::group_by(.data$Location) %>%
  dplyr::summarize_all(sum) %>%
  tibble::column_to_rownames("Location") %>%
  as.matrix

nam <- dimnames(d)
dimnames(d) <- NULL
#d[, 1 : 10]
T_k_ebola <- d[which(nam[[1]] == "GUECKEDOU"), ] # Ebola started in the GUECKEDOU prefecture

# # Observed data (incidence)
print(T_k_ebola)
print(nam[[1]][14])
print(nam[[2]][c(1:3, 71:73)])
barplot(T_k_ebola, xlab = "Week")

K     <- length(T_k_ebola) # number of time periods
t_end <- 7 * K
ts    <- seq(0, t_end, by = 7) # observation schedule

# Observed data
S0 <- 291823 # https://en.wikipedia.org/wiki/Prefectures_of_Guinea
Y_ebola <- list(T_k = T_k_ebola, F_k = NULL, I0 = I0, S0 = S0, ts = ts, t_end = t_end)
```

```{r}
d[which(nam[[1]] == "GUECKEDOU"), ]
d[which(nam[[1]] == "GUECKEDOU"), ][68]
```


```{r}
plot(x = Y_ebola$ts, y = c(0, Y_ebola$T_k), main="Ebola Incidence Observed", sub="each point one month", xlab="Time", ylab="New Infection Counts")
```


# run original constant beta PD-SIR Code

```{r}
#' Experiment 5: Analysis of the Ebola data
#'
#' @inheritParams experiment_1_proof_of_concept
#'
#' @param theta_0 intial parameter values
#' @param path_data path to the data file
#'
#' @return list containing the parameters, observed data, Markov chain and run time of the algorithm
#' @export
experiment_5_ebola <- function(
  I0 = 5, theta_0 = list(R0 = 1, gamma = 1e-1),
  N = 1e5, thin = 10, rho = 1,
  param = "bg", approx = "ldp",
  iota_dist = "exponential",
  gener = FALSE, b = 1/2,
  path_data = "Input/Ebola/Guinea_modified.csv"
) {

  #
  # Ebola Data

  d <- readr::read_csv(
    path_data,
    col_types = readr::cols(
      .default            = readr::col_double(),
      Location            = readr::col_character(),
      `Ebola data source` = readr::col_character(),
      `Indicator type`    = readr::col_character(),
      `Case definition`   = readr::col_character()
    )
  ) %>%
    dplyr::select(-(2:4)) %>%
    dplyr::group_by(.data$Location) %>%
    dplyr::summarize_all(sum) %>%
    tibble::column_to_rownames("Location") %>%
    as.matrix

  nam <- dimnames(d)
  dimnames(d) <- NULL
  #d[, 1 : 10]
  T_k_ebola <- d[which(nam[[1]] == "GUECKEDOU"), ] # Ebola started in the GUECKEDOU prefecture

  K     <- length(T_k_ebola) # number of time periods
  t_end <- 7 * K
  ts    <- seq(0, t_end, by = 7) # observation schedule

  # Observed data
  S0 <- 291823 # https://en.wikipedia.org/wiki/Prefectures_of_Guinea
  Y_ebola <- list(T_k = T_k_ebola, F_k = NULL, I0 = I0, S0 = S0, ts = ts, t_end = t_end)


  MC <- run_DAMCMC(
    Y_ebola, N,
    rho, param, approx,
    iota_dist, gener, b,
    thin, theta_0
  )



  return(list(Y = Y_ebola, MC = MC))

}
```

```{r}
ebola_pdsir <- experiment_5_ebola(
  I0 = 5, theta_0 = list(R0 = 1, gamma = 1e-1),
  N = 20000, thin = 10, rho = 0.2,
  param = "bg", approx = "ldp",
  iota_dist = "exponential",
  gener = FALSE, b = 1/2,
  path_data = "Ebola/Guinea_modified.csv"
)
```

```{r}
ebola_beta <- numeric(length(ebola_pdsir$MC$theta))
for(j in 1:length(ebola_pdsir$MC$theta) ){
  ebola_beta
}
ebola_pdsir$MC$theta[[j]]$beta
```

```{r}
analyze_ebola_pdsir <- experiment_5_output(
  ebola_pdsir, iota_dist = "exponential", path, plot_id = "E5",
  burnin = 0, thin = 1, n_max = NULL
  )
```

```{r}
#' Analyze output of MCMC run from Experiment 5
#'
#' @inheritParams experiment_1_output_analysis
#' @inheritParams experiment_5_ebola
#'
#' @param output_E5 object returned by the function experiment_5_ebola
#'
#' @return figures and list of summary statistics
#' @export
#'
experiment_5_output <- function(
  output_E5, iota_dist = "exponential", path, plot_id = "E5",
  burnin = 0, thin = 1, n_max = NULL
  ) {

  MC <- output_E5[["MC"]]
  Y  <- output_E5[["Y" ]]

  summary <- analyze_MCMC(
    MC, burnin, thin, n_max, iota_dist,
    plot_id = plot_id, path = NULL,
    Y = Y, coverage = FALSE
  )

  return(summary)

}
```


```{r}
original_MC <- run_DAMCMC(
    Y_ebola, N=20000,
    rho=0.2, param = "bg", approx = "ldp",
    iota_dist = "exponential", gener = FALSE, b = 1/2,
    thin = 1, theta_0 = list(R0 = 1, gamma = 1e-1),
    print_i = FALSE,  save_SS = FALSE,
    par_prior = list(
      a_beta = 0.01, b_beta = 1,
      a_gamma = 1, b_gamma = 1,
      a_R0 = 2, b_R0 = 2,
      a_lambda = 1, b_lambda = 1
  ),
  plt = 1000
  )
```

```{r}
original_MC$rate_accept
```

```{r}
# Create dataframe
ebola_constant_beta <- create_mcmc_df(original_MC,1,1000)
```

```{r}
# Posterior Means
ebola_constant_beta %>%
    dplyr::select(- .data$Iteration, - .data$loglik) %>%
    {tibble::tibble(
      var  = colnames(.),
      mean = colMeans(.)
    )}

ESS         = coda::effectiveSize(coda::mcmc(ebola_constant_beta))
ESS
``` 

```{r}
# Plot constant beta results
print(draw_tp(ebola_constant_beta, "beta"))
print(draw_histogram(ebola_constant_beta, "beta"))
print(draw_acf(ebola_constant_beta, "beta", 200))
```


# Change Point D-A MCMC

# Set up condensed version of ebola data.
```{r}
T_k_ebola_short <- c()
for(j in seq(from=1, to=length(T_k_ebola), by=5)){
  # print(j)
  right <- j+4
  # print(cut)
  T_k_ebola_short <- append(T_k_ebola_short, sum(T_k_ebola[j:right]))
  # print(T_k_ebola_short)
}
T_k_ebola_short
T_k_ebola_short <- T_k_ebola_short[1:12]
```

```{r}
K     <- length(T_k_ebola_short) # number of time periods

Y_ebola_cp <- list(T_k = T_k_ebola_short, F_k = NULL, I0 = 5, S0 = 291823, ts = seq(0, 12, by = 1), t_end = 12)
```

```{r}
plot(x = Y_ebola_cp$ts, y = c(0, Y_ebola_cp$T_k), main="Ebola Incidence Observed", sub="each point one month", xlab="Time", ylab="New Infection Counts")
```


```{r}
ebola_theta0 <- list("R0" = 1, "beta" = 3e-7, "gamma" = 1e-1, "shape"=1, "lambda"=1)
```

# Run change point algorithm
```{r}
ebola_output <- run_DAMCMC_complete_2(
  ebola_theta0, Y_ebola_cp, N = 20000,
  rho = 0.2, param = "bg", approx = "ldp",
  iota_dist = "exponential", gener = FALSE, b = 1/2,
  thin = 1, plt = 500,
  par_prior = list(
    a_beta = 0.01, b_beta = 0.01,
    a_gamma = 1, b_gamma = 1,
    a_R0 = 2, b_R0 = 2,
    a_lambda = 1, b_lambda = 1
  ), h_hat = c(0.50,0.50,0.50,0.50,0.50,0.50,0.50,0.50,0.50,0.50,0.0,0.0), # P(dbeta_i = 1)
  x_b_ratio = 2,
  gamma=0.1
)
```

```{r}
burnin <- 501
mcmc.trace.ebola <- create_tp_data(ebola_output$beta_lst[burnin:20000])
summary(mcmc.trace.ebola)
```

```{r}
plot(mcmc.trace.ebola)
```

```{r}
ebola_output$theta_rate_accept # theta = (dbeta, beta).
ebola_output$x_rate_accept # latent data.
```


```{r}
analyze_dbeta(ebola_output$dbeta_lst)[[1]]
unique(ebola_output$dbeta_lst) # frequencies of each.
```

# Create Dataframe from raw dbeta and beta lists for frequency plot of configurations.
```{r}
create_df_dbeta <- function(mcmc_output_dbeta, mcmc_output_beta, burnin, strlastX) {
  # strlastX: string value of the last dbeta (change point) index.
  temp_df <- mcmc_output_dbeta[burnin:length(mcmc_output_dbeta)]
  temp_df <- data.frame(t(sapply(temp_df,c)))
  dbeta_lst_df <- temp_df %>% 
    mutate(ID = burnin:length(mcmc_output_dbeta)) %>% 
    unite("delta_beta", X1:strlastX, remove=FALSE)

  beta_lst_raw <- mcmc_output_beta[burnin:length(mcmc_output_beta)]
  beta_lst_df <- data.frame(t(sapply(beta_lst_raw,c)))
  beta_lst_df <- beta_lst_df %>% 
    mutate(ID = burnin:length(mcmc_output_beta))
  
  complete_tibble <- merge(dbeta_lst_df, beta_lst_df, by="ID")

  return(complete_tibble)
}

config_df <- create_df_dbeta(ebola_output$dbeta_lst, ebola_output$beta_lst, burnin, "X12")
```

# Frequencies of the most common configurations
```{r}
config_df %>% 
  count(delta_beta) %>% 
  mutate(freq = n / sum(n)) %>% 
  arrange(desc(n)) %>% 
  head(10) %>% 
  ggplot(data = ., aes(x = delta_beta, y = freq)) +
  geom_bar(stat = "identity") +
  labs(title = "Frequency of Configurations", y = "Frequency", x = "Configuration")+ 
  theme(text = element_text(size=10), axis.text.x = element_text(angle = 90))
```
```{r}
top_five <- config_df %>% 
  count(delta_beta) %>% 
  mutate(freq = n / sum(n)) %>% 
  arrange(desc(n)) %>% 
  head(5)
top_five
```

# Compute the mean betas of the top configurations.
```{r}
top_deltabetas <- config_df %>% 
  filter(delta_beta %in% top_five$delta_beta)

mean_betas <- top_deltabetas %>%
  select(delta_beta, X1.y:X13) %>%
  group_by(delta_beta) %>% 
  summarise(across(everything(), mean))

# flip from cartesian to index. https://tidyr.tidyverse.org/reference/pivot_longer.html
M <- mean_betas %>% 
  pivot_longer(!delta_beta, names_to = "time_unit", values_to = "beta")

#preserve the order of time units.
M$time_unit <- factor(M$time_unit, levels=unique(M$time_unit))
```

```{r}
# Colors
cols <- c("grey", "orange", "red", "yellow", "grey")

# Basic line plot with points
ggplot(data=M, aes(x=time_unit, y=beta, group=delta_beta, color = delta_beta)) +
  geom_step() +
  # geom_line()+
  geom_point() +
  scale_color_manual(values = cols) +
  labs(title = "Transmission Rate of Ebola Outbreak", subtitle="Mean of Top 5 Configurations", y = "Beta value", x = "Time unit")

```
# top three # 1: 0_0_1_0_1_0_0_1_0_1_0_0, 2: 0_0_1_0_1_0_0_1_0_0_0_0, 3: 0_0_1_0_1_0_0_1_1_0_0_0

# spagetti plot of the overall (every 100 iterations)


```{r}
plot(x = Y_ebola_cp$ts, y = c(0, Y_ebola_cp$T_k), main="Ebola Incidence Observed", sub="each point one month", xlab="Time", ylab="New Infection Counts")
abline(v=c(3,5,8,10),lwd=2,col="red") # top configuration
```



```{r}
analyze_dbeta(ebola_output$dbeta_lst)

barplot(analyze_dbeta(ebola_output$dbeta_lst)[[2]], main="Delta Configurations", xlab="Change Point Location",  
 ylab="Count", names.arg=c("Interval 1","Interval 2","Interval 3","Interval 4","Interval 5", "Interval 6","Interval 7","Interval 8","Interval 9","Interval 10","Interval 11","Interval 12"), 
 border="blue")
abline(h=0.5, col="blue", lty="dashed", lwd=4)
```

# Create dataframe
```{r}
create_mcmc_list <- function(beta_list_untidy){
  # make each beta into a plain list
  all_beta <- list()
  
  for(k in 1:length(beta_list_untidy[[1]])){
    index <- paste("beta", k, sep="_")
    beta_i = unlist(lapply(beta_list_untidy, `[[`, k)) # beta k
    all_beta <- append(all_beta, list(beta_i))
  }

  return(all_beta)
}

beta_list_ebola <- create_mcmc_list(ebola_output$beta_lst[burnin:20000])
length(beta_list_ebola)
```

```{r}
# combine all beta lists into a dataframe.
beta_ebola_tidy <- do.call(rbind, Map(data.frame, Iteration = list(c(1:19500)), Beta1 = beta_list_ebola[1], Beta2 = beta_list_ebola[2], Beta3 = beta_list_ebola[3], Beta4 = beta_list_ebola[4], Beta5 = beta_list_ebola[5], Beta6 = beta_list_ebola[6], Beta7 = beta_list_ebola[7],Beta8 = beta_list_ebola[8], Beta9 = beta_list_ebola[9], Beta10 = beta_list_ebola[10], Beta11 = beta_list_ebola[11], Beta12 = beta_list_ebola[12], Beta13 = beta_list_ebola[13]))
beta_ebola_tidy <- beta_ebola_tidy %>% 
  remove_burnin(1000)
```

```{r}
# Posterior means
beta_ebola_tidy %>%
  dplyr::select(- .data$Iteration) %>%
  {tibble::tibble(
    var  = colnames(.),
    mean = colMeans(.)
  )}

# ESS
ESS = coda::effectiveSize(coda::mcmc(beta_ebola_tidy))
ESS
```

```{r}
# plot results in a format similar to PDSIR code.
vars <- c("Beta1", "Beta2","Beta3","Beta4","Beta5","Beta6")

for(var in vars)
  print(draw_tp(beta_ebola_tidy, var))
for(var in vars)
  print(draw_histogram(beta_ebola_tidy, var))
for(var in vars)
  print(draw_acf(beta_ebola_tidy, var, 200))
```
Why are plots mixing well in the change point algorithm? Does it have to do with keepign a constant beta?

# Finer grained time series.
```{r}
Y_ebola$ts
length(Y_ebola$ts)

h1 <- rep(0.5, 5)
h2 <- rep(0.5, 47)
h3 <- rep(0.5, 21)

h <- c(h1, h2, h3)
h
length(h)
```

```{r}
Y_cp$ts <- seq(0,73,1)
Y_cp
```


```{r}
ebola_cp <- run_DAMCMC_complete(
  ebola_theta0, Y_cp, N = 20000,
  rho = 0.2, param = "bg", approx = "ldp",
  iota_dist = "exponential", gener = FALSE, b = 1/2,
  thin = 1, plt = 500,
  par_prior = list(
    a_beta = 1, b_beta = 1,
    a_gamma = 1, b_gamma = 1,
    a_R0 = 2, b_R0 = 2,
    a_lambda = 1, b_lambda = 1
  ), h, # P(dbeta_i = 1)
  x_b_ratio = 2,
  gamma=0.8
)
```


```{r}
analyze_dbeta(ebola_cp$dbeta_lst)[[1]]
length(unique(ebola_cp$dbeta_lst)) # frequencies of each.
unique(ebola_cp$dbeta_lst)[180:188]
```

```{r}
analyze_dbeta(ebola_beta_5$dbeta_lst)[[1]]
unique(ebola_beta_5$dbeta_lst)[300:310]
```


# Create Dataframe from raw dbeta and beta lists for frequency plot of configurations.
```{r}
config_df <- create_df_dbeta(ebola_cp$dbeta_lst, ebola_cp$beta_lst, burnin, "X73")
```
# Frequencies of the most common configurations
```{r fig.width=5, fig.height=10}
config_df %>% 
  count(delta_beta) %>% 
  mutate(freq = n / sum(n)) %>% 
  arrange(desc(n)) %>% 
  head(10) %>% 
  ggplot(data = ., aes(x = delta_beta, y = freq)) +
  geom_bar(stat = "identity") +
  labs(title = "Frequency of Configurations", y = "Frequency", x = "Configuration")+ 
  theme(text = element_text(size=5), axis.text.x = element_text(angle = 90))
```

```{r}
top_five <- config_df %>% 
  count(delta_beta) %>% 
  mutate(freq = n / sum(n)) %>% 
  arrange(desc(n)) %>% 
  head(5)
top_five
```

# Compute the mean betas of the top configurations.
```{r}
top_deltabetas <- config_df %>% 
  filter(delta_beta %in% top_five$delta_beta)

mean_betas <- top_deltabetas %>%
  select(delta_beta, X1.y:X60.y) %>%
  group_by(delta_beta) %>% 
  summarise(across(everything(), mean))

# flip from cartesian to index. https://tidyr.tidyverse.org/reference/pivot_longer.html
M <- mean_betas %>% 
  pivot_longer(!delta_beta, names_to = "time_unit", values_to = "beta")

#preserve the order of time units.
M$time_unit <- factor(M$time_unit, levels=unique(M$time_unit))
```

```{r fig.width=6, fig.height=5}
# Color selection
cols <- c("red", "orange", "yellow", "green", "blue")

# Basic line plot with points
ggplot(data=M, aes(x=time_unit, y=beta, group=delta_beta, color = delta_beta)) +
  # geom_step() +
  geom_line()+
  geom_point() +
  scale_color_manual(values = cols) +
  labs(title = "Transmission Rate of Ebola Outbreak", subtitle="Mean of Top 5 Configurations", y = "Beta value", x = "Time unit")+ 
  theme(legend.position="none") +
  theme(text = element_text(size=10), axis.text.x = element_text(angle = 90))
```


```{r}
analyze_dbeta(ebola_cp$dbeta_lst)

barplot(analyze_dbeta(ebola_cp$dbeta_lst)[[2]], main="Delta Configurations", xlab="Change Point Location",  
 ylab="Count", 
 border="blue")
abline(h=0.5, col="blue", lty="dashed", lwd=4) # change points around 15,40,71
```


```{r}
ebola_cp$theta_rate_accept
ebola_cp$x_rate_accept
```


```{r}
burnin <- 501
mcmc.trace.ebolacp <- create_tp_data(ebola_cp$beta_lst[burnin:20000])
summary(mcmc.trace.ebolacp)
```

```{r}
plot(x = Y_cp$ts[-73]+0.5, y = c(Y_cp$T_k), main="Ebola", sub="week", xlab="Time", ylab="New Infection Counts")
# abline(v=c(15,41,52),lwd=2,col="red") # dots should be 0.5 shifted.
abline(v=c(15,40,71),lwd=2,col="red")
```


```{r}
ebola_cp_gamma <- run_DAMCMC_complete_bug(
  ebola_theta0, Y_cp, N = 20000,
  rho = 0.2, param = "bg", approx = "ldp",
  iota_dist = "exponential", gener = FALSE, b = 1/2,
  thin = 1, plt = 500,
  par_prior = list(
    a_beta = 1, b_beta = 1,
    a_gamma = 1, b_gamma = 1,
    a_R0 = 2, b_R0 = 2,
    a_lambda = 1, b_lambda = 1
  ), h, # P(dbeta_i = 1)
  x_b_ratio = 2,
  gamma=1
)
```

```{r}
analyze_dbeta(ebola_cp_gamma$dbeta_lst)

barplot(analyze_dbeta(ebola_cp_gamma$dbeta_lst)[[2]], main="Delta Configurations", xlab="Change Point Location",  
 ylab="Count", 
 border="blue")
abline(h=0.5, col="blue", lty="dashed", lwd=4)
```
# intervals too short: additional change points due to 


```{r}
ebola_cp_gamma <- run_DAMCMC_complete_bug(
  ebola_theta0, Y_cp, N = 20000,
  rho = 0.2, param = "bg", approx = "ldp",
  iota_dist = "exponential", gener = FALSE, b = 1/2,
  thin = 1, plt = 500,
  par_prior = list(
    a_beta = .01, b_beta = .01,
    a_gamma = 1, b_gamma = 1,
    a_R0 = 2, b_R0 = 2,
    a_lambda = 1, b_lambda = 1
  ), h, # P(dbeta_i = 1)
  x_b_ratio = 2,
  gamma=1
)
```

